---
title: "InsightFlow — Executive Business Report"
author: "InsightFlow Analytics Platform"
date: today
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    colorlinks: true
    mainfont: "Inter"
    fontsize: 11pt
    geometry: "margin=1in"
    header-includes:
      - \usepackage{booktabs}
      - \usepackage{graphicx}
params:
  db_path: "insightflow.db"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 7, fig.height = 3.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(DBI)
library(RSQLite)

# Connect to database
con <- dbConnect(SQLite(), params$db_path)
sales <- dbReadTable(con, "sales") %>% mutate(date = as.Date(date))
expenses <- dbReadTable(con, "expenses") %>% mutate(date = as.Date(date))
customers <- dbReadTable(con, "customers")
inventory <- dbReadTable(con, "inventory")
dbDisconnect(con)
```

# Executive Summary

This report provides a comprehensive analysis of business performance generated by **InsightFlow**, an AI-powered analytics platform.

```{r kpi-table}
total_revenue <- sum(sales$total, na.rm = TRUE)
total_expenses <- sum(expenses$amount, na.rm = TRUE)
gross_profit <- total_revenue - total_expenses
margin <- if (total_revenue > 0) (gross_profit / total_revenue) * 100 else 0

kpi_df <- data.frame(
  Metric = c("Total Revenue", "Total Expenses", "Gross Profit",
             "Gross Margin", "Total Orders", "Total Customers"),
  Value = c(
    sprintf("$%s", format(round(total_revenue), big.mark = ",")),
    sprintf("$%s", format(round(total_expenses), big.mark = ",")),
    sprintf("$%s", format(round(gross_profit), big.mark = ",")),
    sprintf("%.1f%%", margin),
    format(nrow(sales), big.mark = ","),
    format(nrow(customers), big.mark = ",")
  ),
  stringsAsFactors = FALSE
)

kable(kpi_df, booktabs = TRUE, align = "lr",
      caption = "Key Performance Indicators")
```

# Sales Analysis

## Monthly Revenue Trend

```{r revenue-trend, fig.height=3}
monthly <- sales %>%
  mutate(month = as.Date(format(date, "%Y-%m-01"))) %>%
  group_by(month) %>%
  summarise(revenue = sum(total, na.rm = TRUE), .groups = "drop")

ggplot(monthly, aes(x = month, y = revenue)) +
  geom_area(fill = "#6366f1", alpha = 0.1) +
  geom_line(color = "#6366f1", linewidth = 1) +
  geom_point(color = "#4f46e5", size = 2) +
  scale_y_continuous(labels = dollar_format()) +
  scale_x_date(date_labels = "%b %Y") +
  labs(x = NULL, y = "Revenue") +
  theme_minimal(base_size = 10) +
  theme(panel.grid.minor = element_blank())
```

## Top 10 Products by Revenue

```{r top-products}
top <- sales %>%
  group_by(Product = product) %>%
  summarise(Revenue = sum(total, na.rm = TRUE),
            Orders = n(),
            Units = sum(quantity, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(desc(Revenue)) %>%
  head(10) %>%
  mutate(Revenue = sprintf("$%s", format(round(Revenue), big.mark = ",")))

kable(top, booktabs = TRUE, align = "lrrr",
      caption = "Top 10 Products by Revenue")
```

## Revenue by Category

```{r category-chart, fig.height=3}
cats <- sales %>%
  group_by(category) %>%
  summarise(revenue = sum(total, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(revenue))

ggplot(cats, aes(x = reorder(category, revenue), y = revenue)) +
  geom_col(fill = "#06b6d4") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = NULL, y = "Revenue") +
  theme_minimal(base_size = 10)
```

# Expense Analysis

```{r expense-table}
exp_cat <- expenses %>%
  group_by(Category = category) %>%
  summarise(Total = sum(amount, na.rm = TRUE),
            Count = n(), .groups = "drop") %>%
  arrange(desc(Total)) %>%
  mutate(Total = sprintf("$%s", format(round(Total), big.mark = ",")))

kable(exp_cat, booktabs = TRUE, align = "lrr",
      caption = "Expenses by Category")
```

# Inventory Status

```{r inventory-status}
if (nrow(inventory) > 0) {
  inv_status <- inventory %>%
    mutate(
      Status = case_when(
        stock_quantity <= reorder_level * 0.5 ~ "CRITICAL",
        stock_quantity <= reorder_level ~ "LOW",
        TRUE ~ "OK"
      )
    )

  low <- inv_status %>% filter(Status != "OK")

  if (nrow(low) > 0) {
    cat(sprintf("**%d items** require attention:\n\n", nrow(low)))
    kable(low %>% select(Product = product, Stock = stock_quantity,
                          `Reorder Level` = reorder_level, Status),
          booktabs = TRUE, align = "lrrl",
          caption = "Items at or Below Reorder Level")
  } else {
    cat("All inventory levels are healthy.\n")
  }
}
```

# Recommendations

Based on the automated analysis, InsightFlow recommends:

1. **Revenue Growth**: Focus marketing efforts on top-performing categories to maximize ROI.
2. **Inventory Management**: Reorder items flagged as LOW or CRITICAL to avoid stockouts.
3. **Cost Optimization**: Review the highest expense categories for potential savings.
4. **Customer Retention**: Implement targeted campaigns for at-risk customer segments.
5. **Forecasting**: Use the built-in forecasting engine to project demand and plan inventory.

---

*This report was automatically generated by InsightFlow v1.0.0*
